Viscous Transport in Eroding Porous Media

### 2019-05-14 docs/transport/erosion_transport.tex
changes to section 2 of transport paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
Continuing with our previous work~\cite{qua-moo2018}, we use a completed
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, $\nn$ is the unit outward
normal at $\yy$, and $\eeta$ is an unknown density function.  The Stokes
double-layer potential can also be expressed as the sum of a Laplace
double-layer potential 
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy
\end{align}
and its gradient.  By identifying $\xx,\yy \in
\RR^2$ with the points $x,y \in \CC$, the Laplace double-layer potential
for a real-valued density function $\eta$ can also be written as the
real part of the Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy.
\end{align}
Therefore, the Stokes double-layer potential can be written in terms of
Cauchy integrals and its first derivative.  Rather than immediately
applying a quadrature rule to compute $v(x)$, we will first compute the
boundary data of $v(x)$ by applying the Sokhotski-Plemelj jump relation
\begin{align}
  v^{-}(x) = -\frac{1}{2} \eta(x) - \frac{1}{2\pi i} \text{PV}
\end{align}

In addition, we also require
expressions for the derivative of the velocity field which can be
written in terms of a Cauchy integral and its first two derivatives.
Therefore, we introduce the notation
\todo[inline]{BQ: Need to discuss boundary data and how to get it from
density $\eta$}
\begin{align}
  \label{eqn:cauchy}
  v[\eta](x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,  \quad
  v'[\eta](x) = -\frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^2} \, dy, \quad
  v''[\eta](x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^3} \, dy.
\end{align}
To simplify notation, whenever it will not cause confusion, we exclude
$\tau$ and instead write $v(x)$, $v'(x)$, and $v''(x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
The Stokes double-layer potential~\eqref{eqn:stokesDLP} can be
decomposed as
\begin{align}
  \DDD[\eeta](\xx) &= 
    \frac{1}{2\pi} \int_{\bd\Omega} 
      \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
    \frac{1}{2\pi} \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
    &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
    \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
\end{align}
All these integrals but the first are a gradient of a Laplace layer
potential.  In addition, each component of the first integral can be
written as a Cauchy integral with a complex-valued density
function~\cite{bar-wu-vee2015}.  Therefore, the Stokes double-layer
potential can be written as five Cauchy integrals and their derivatives
with the density functions $\yy \cdot \eeta$, $\eta_1$, and $\eta_2$.
The individual velocity components are
\begin{align}
  \label{eqn:cauchy1}
  u_1(\xx) &= \Re (v[\tau_1](x)) + \Re (v'[\yy\cdot\eeta](x)) 
           -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
  u_2(\xx) &= \Re (v[\tau_2](x)) - \Im (v'[\yy\cdot\eeta](x)) 
       +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
\end{align}
where
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Shear Stress}
To compute the rate of erosion, we must compute the shear stress of the
double-layer potential
\begin{align}
  \tau(\xx) = \left(\nabla \DDD[\eeta](\xx) + 
    \nabla^T \DDD[\eeta](\xx) \right) \nn \cdot \ss,
\end{align} where $\nn$ and $\ss$ are the unit normal and tangent
vectors at $\xx$, respectively.  Each term of the gradient of the
velocity field can be written in terms of Cauchy integrals.  For
example,
\begin{align}
  \pderiv{u_1}{x_2} &= \Re(iv'[\tau_1](x)) + 
      \Re(i v''[\yy \cdot \nn](x)) -
      x_1\Re(i v''[\eta_1](x))  -
      \Re(v'[\eta_2](x)) - x_2 \Re(iv''[\eta_2](x)) \\
    &= -\Im(v''[\tau_1](x)) - 
      \Im(v''[\yy \cdot \nn](x)) +
      x_1\Im(v''[\eta_1](x)) -
      \Re(v'[\eta_2](x)) + x_2 \Im(v''[\eta_2](x)).
\end{align}
Similarly, the other three components of $\nabla \DDD[\eeta]$ can be
calculated
\begin{align}
  \pderiv{u_1}{x_1} &= \Re (v'[\tau_1](x)) + 
    \Re (v''[\yy\cdot\eeta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[\yy\cdot\eeta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= \Re (v'[\tau_2](x)) - 
    \Im (v''[\yy\cdot\eeta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[\yy\cdot\eeta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense1}
\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense101}\\

\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense201}
\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies with the fixed-pressure-drop condition. The color is the vorticity of fluid. The 4 snapshots are evenly spaced in time. In this simulation, we use $N_\iin = 256$ discretization points on the body, $N_\out = 1024$ points on the outer boundary, a time-step of $\Dt=10^{-4}$, and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $. In the forth frame, bodies 4, 12, and 17 have vanished completely, and body 19 (the one without number on the right) has almost vanished because of eroding.}
\end{center}
\end{figure}
\subsection{20 nearly touching bodies}
{\color{red} 
By using Barycentric quadrature rule, we are able to capture the details in the fluid when wall and bodies or body and bodies are close to each other. The first nearly touching bodies case is 20 bodies in a Hagen-Poiseuille flow. The domain $\Omega$ is a rectangle $[-3,3]\times[-1,1]$ with rounded corners. The velocity of flow from the left side of $\Omega$ is $v(x_1,x_2)=(1+x_2)(1-x_2)$. The number of discretization points $N_\iin$ on the bodies is 256 and the number of discretization points $N_\out$ on the outer boundary is 1024.

The snapshots of the process of 20 nearly touching bodies eroding are in Fig.~\ref{fig:Eroding20vort}. 
We place some bodies to nearly contact not only one another but also the outer boundary. For example, bodies 1, 6, 13, and 15 are very close to outer boundary. In~\cite{qua-moo2018} we reported the quick expanding space at the narrow channel between neighboring bodies, the developing flat faces at the sites of near contact, and bodies vanishing. In this nearly touching bodies structure, we also observe these phenomena but with more new details.
For the speed of expanding space at the narrow channel, we see some narrow channels do expand faster than other wider channels. For example, the spacing between bodies 14 \& 18 is smaller than the spacing between bodies
 7 \& 14 initially and the former one is comparable to the latter one by the second frame. But there are some small initial spacings that do not expand that much at all by the fourth frame, for example between bodies 1 \& 6, 3 \& 9. 
We could find the explanation from the trajectories of tracers in Fig.~\ref{fig:Eroding20tracer}. Since the direction of flow is from left to right, most of the tracers will not travel vertically but horizontally when the gaps or channels are too small.  }

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
Continuing with our previous work~\cite{qua-moo2018}, we use a
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, $\nn$ is the unit outward
normal at $\yy$, and $\eeta$ is an unknown density function.  In
Section~\ref{sec:DLPcomplex}, we show how the Stokes double-layer
potential can also be expressed as the sum of a Laplacian double-layer
potential 
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy
\end{align}
and its gradient.  Since the Laplacian double-layer potential can be
written as a Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \Re (v(x)) \quad \text{where} \quad
  v(x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,
  \label{eqn:laplaceComplex}
\end{align}
where $\xx,\yy \in \RR^2$ are identified with the points $x,y \in \CC$,
the Stokes double-layer potential can be written in terms of Cauchy
integrals and its first derivative.  We also require the derivative of
the Stokes double-layer potential which requires the second derivative
of Cauchy integrals.

We will be using quadrature formulae to compute $v$ and its derivatives
that rely on the Cauchy integral theorem.  Since $v$ is holomorphic, its
value only depends on its boundary data.  The boundary data satisfies
the Sokhotski-Plemelj jump relation
\begin{align}
  v(x) = -\frac{1}{2} \eta(x) + \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \bd\Omega.
\end{align}
Once the boundary data of $v$ is computed, we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = -\frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy.
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the density function $\eta
\in \CC$, we use the notation $v[\eta](x)$ to denote the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and the
derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
The Stokes double-layer potential~\eqref{eqn:stokesDLP} can be
decomposed as
\begin{align}
  \DDD[\eeta](\xx) &= 
    \frac{1}{2\pi} \int_{\bd\Omega} 
      \frac{\nn}{\rho^2} (\rr \cdot \eeta) \, ds_\yy + 
    \frac{1}{2\pi} \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2} (\yy \cdot \eeta) \, ds_\yy \\
    &- \frac{1}{2\pi} x_1 \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2}\eta_1(\yy) \, ds_\yy -
    \frac{1}{2\pi} x_2 \nabla \int_{\bd\Omega}
      \frac{\rr \cdot \nn}{\rho^2}\eta_2(\yy) \, ds_\yy.
\end{align}
All these integrals but the first are a gradient of a Laplace layer
potential.  In addition, each component of the first integral can be
written as a Cauchy integral with a complex-valued density
function~\cite{bar-wu-vee2015}.  Therefore, the Stokes double-layer
potential can be written as five Cauchy integrals and their derivatives
with the density functions $\yy \cdot \eeta$, $\eta_1$, and $\eta_2$.
The individual velocity components are
\begin{align}
  \label{eqn:cauchy1}
  u_1(x) &= \Re (v[\tau_1](x)) + \Re (v'[y\cdot\eta](x)) 
           -x_1\Re (v'[\eta_1](x)) - x_2\Re (v'[\eta_2](x)), \\
  u_2(x) &= \Re (v[\tau_2](x)) - \Im (v'[y\cdot\eta](x)) 
       +x_1\Im (v'[\eta_1](x)) + x_2\Im (v'[\eta_2](x)),
\end{align}
where $y \cdot \eta = y_1 \eta_1 + i y_2 \eta_2 \in \CC$, 
\begin{align} 
  \tau_1=(\eta_1+i\eta_2)\frac{\Re(n)}{n}, \quad \text{and} \quad
  \tau_2=(\eta_1+i\eta_2)\frac{\Im(n)}{n},
\end{align}
and $n \in \CC$ is the complex representation of the outward unit normal
$\nn \in \RR^2$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Shear Stress}
To compute the rate of erosion, we must compute the shear stress of the
double-layer potential
\begin{align}
  \tau(\xx) = \left(\nabla \DDD[\eeta](\xx) + 
    \nabla^T \DDD[\eeta](\xx) \right) \nn \cdot \ss,
\end{align} where $\nn$ and $\ss$ are the unit normal and tangent
vectors at $\xx$, respectively.  Each term of the gradient of the
velocity field can be written in terms of Cauchy integrals.  For
example,
\begin{align}
  \pderiv{u_1}{x_2} &= \Re(iv'[\tau_1](x)) + 
      \Re(i v''[y \cdot \eta](x)) -
      x_1\Re(i v''[\eta_1](x))  -
      \Re(v'[\eta_2](x)) - x_2 \Re(iv''[\eta_2](x)) \\
    &= -\Im(v''[\tau_1](x)) - 
      \Im(v''[y \cdot \eta](x)) +
      x_1\Im(v''[\eta_1](x)) -
      \Re(v'[\eta_2](x)) + x_2 \Im(v''[\eta_2](x)).
\end{align}
Similarly, the other three components of $\nabla \DDD[\eeta]$ can be
calculated
\begin{align}
  \pderiv{u_1}{x_1} &= \Re (v'[\tau_1](x)) + 
    \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= \Re (v'[\tau_2](x)) - 
    \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense_1}
\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense_101}\\

\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense_201}
\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense_301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies with the fixed-pressure-drop condition. The color is the vorticity of fluid. The 4 snapshots are evenly spaced in time. In this simulation, we use $N_\iin = 256$ discretization points on the body, $N_\out = 1024$ points on the outer boundary, a time-step of $\Dt=10^{-4}$, and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $.}
\end{center}
\end{figure}
\subsection{20 nearly touching bodies}
{\color{red} 
By using Barycentric quadrature rule, we are able to capture the details in the fluid when wall and bodies or body and bodies are close to each other. The first nearly touching bodies case is 20 bodies in a Hagen-Poiseuille flow. The domain $\Omega$ is a rectangle $[-3,3]\times[-1,1]$ with rounded corners. The velocity of flow from the left side of $\Omega$ is $v(x_1,x_2)=(1+x_2)(1-x_2)$. The number of discretization points $N_\iin$ on the bodies is 256 and the number of discretization points $N_\out$ on the outer boundary is 1024.

The snapshots of the process of 20 nearly touching bodies eroding are in Fig.~\ref{fig:Eroding20vort}. 
We place some bodies to nearly contact not only one another but also the outer boundary. For example, bodies 1, 6, 13, and 15 are very close to outer boundary. In~\cite{qua-moo2018} we reported the quick expanding space at the narrow channel between neighboring bodies, the developing flat faces at the sites of near contact, and bodies vanishing. In this nearly touching bodies structure, we also observe these phenomena but with more new details.
For the speed of expanding space at the narrow channel, we see some narrow channels do expand faster than other wider channels. For example, the spacing between bodies 14 \& 18 is smaller than the spacing between bodies
 7 \& 14 initially and the former one is comparable to the latter one by the second frame. But there are some small initial spacings that do not expand that much at all by the fourth frame, for example between bodies 1 \& 6, 3 \& 9. 
We could find the explanation from the tracer test in Fig.~\ref{fig:Eroding20tracer}. Since the direction of flow is from left to right, most of the tracers will not travel vertically but horizontally. If the trajectory of tracer }
